---
title: "Week 4-5 R Lab"
author: "Gene Ho"
date: "4/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE)
```

For Questions 1 and 3, skeleton code is given. To complete the lab, update the code to correctly execute the specifications of the question. 

# Question 1

Install the packages listed in the lab handout: 
"rmarkdown", "knitr", "epiR", "stats", "tidyverse", "readxl", "epitools", "exact2x2", "DescTools", "MESS", "pwr", "here"

Remember to quote the package name when running `install.packages` since you're technically searching the internet for the package name. 

```{r}
install.packages('rmarkdown')
# install the rest of the packages!
```

Change the code below to load the `tidyverse` package. Note, you don't have to quote the package name here since R knows to osearch your finite selection of installed packages. 

```{r}
# this cell + output goes in the lab (once updated)
library(rmarkdown)
```

# Question 2 

Run the code below to give us information on your local RStudio installation. 

```{r}
Sys.info()

R.version
```


# Question 3

Import the `inorg_p.xlsx` file and run `str`. You may need to load a certain package to get this function to run. If you're not sure what the package is called, run "??function_name_here". Remember that R sometimes lists functions as "package_name::function_name" (see pages 4-5 of RStudio Basics for more info). 

```{r}
whatever_name_you_want_to_call_this_obj <- read_excel('path_to_file/inorg_p.xlsx') 
# might be in the data folder

str(obj_name_created_in_line_above)
```

