---
title: "pre-schooler's spatio-temporal reasoning and piano lessons"
author: "Maureen Lahiff"
date: "March 22, 2019"

---

# Moore and McCabe's preschooler's piano lesson data n = 34

# data file is piano_reasoning.xlsx
# variable (column) name reas_diff

# working with confidence intervals for means
# practice with box plots and qqnorm plots to check assumption
# for small samples, 
# so we know sampling dist of x-bar is approx normal curve

# then we'll practice with 95% and 99% confidence intervals

# the code in this R Markdown file is a template for you to adapt
# for other practice and the problem set

```{r the usual directory and libraries, plus a new one: epiR}

# be sure to set your working directory or project

library(tidyverse)
library(readr)
library(readxl)
library(epiR)

```


```{r read in data for reas_diff}

piano_reas <- read_excel("piano_reasoning.xlsx", col_names = TRUE)

```

```{r summary statistics and boxplot}

summary(piano_reas$reas_diff)

piano_reas %>% summarize(
    min = min(reas_diff) ,
    q1 = sprintf("%0.2f", quantile(reas_diff, 0.25)) ,
    median = sprintf("%0.2f", quantile(reas_diff, 0.50)) ,
    q3 = sprintf("%0.2f", quantile(reas_diff, 0.75)) ,
    max = max(reas_diff)
)

ggplot(piano_reas, aes(y = reas_diff) ) +
  geom_boxplot(width = 0.5) +
    scale_x_continuous(labels = NULL) +
    labs(y = "difference in reasoning scores",       
         x = "", title = "preschooler's spatio-temoporal reasoning", subtitle = " differences after 6 months of piano lessons, n = 34") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))

```

```{r normal quantile plot}

qqnorm(piano_reas$reas_diff, frame = FALSE,
       main = "normal quantile plot reasoning score diffs",
       xlab = "normal scores", ylab = "differences")
qqline(piano_reas$reas_diff)

```

```{r confidence intervals}

# calculate and save mean and se of the diffs
mean_diff = mean(piano_reas$reas_diff)
std_dev_diff = sd(piano_reas$reas_diff)
se_mean_diff <- sd(piano_reas$reas_diff) / sqrt(34)

# get t* value for 95% confidence interval
t_star <- qt(.975, df = 33, lower.tail = TRUE)

#calculate confidence interval 
lower_limit <- mean_diff - t_star * se_mean_diff
upper_limit <- mean_diff + t_star * se_mean_diff

# print mean, se, and confidence interval
c(mean_diff, std_dev_diff, se_mean_diff)
c(lower_limit, upper_limit)

# use epi.conf from epiR to calculate 95% confidence interval
epi.conf(piano_reas$reas_diff, ctype = "mean.single", conf.level = 0.95)

# use epi.conf from epiR to calculate 99% confidence interval
epi.conf(piano_reas$reas_diff, ctype = "mean.single", conf.level = 0.99)

```

```{r Shapiro-Wilk test for Week 5}

shapiro.test(piano_reas$reas_diff)

```

```{r t test for Week 5}

t.test(piano_reas$reas_diff, mu = 0, alternative = "greater", conf.level = .95)

```

