---
title: "bar charts for categorical variables"
author: "Maureen Lahiff"
date: "February 21, 2020"
output: word_document
---

```{r}
source(here::here('0-setup.R'))

# read in the data
nhanes <- readxl::read_xlsx('nhanes_adult_1516.xlsx') %>% 
        # cleaning the data
        # see build_cat_tables.Rmd for more detail
        mutate(gender_factor = factor(gender, 
                                      labels = c('male', 'female')), 
               age_cat_factor = factor(educ_level, 
                                       labels = c("lt hs", "some hs", "hs grad",
                                                  "some coll", "coll grad")),
               ed_lvl_factor = factor(age_cat,
                                      labels = c("20 to 34","35 to 49",
                                                 "50 to 64","65 and older")) )
        
```


# the table commands handled missing observations gracefully
# in the Global Environment, you can see that nhanes has 5719 observations
# however, the total for the tables we made with education level is 5714 obervations

# many, if not most, R functions are not so nice

# ggplot will create a separate category for the 5 missing values of rel_stat
# which is unlikely to be what we want

# to deal with this, create a new data frame with these rows removed

```{r create a smaller data frame and then remove rows with missing values}

educ_age_gender <- select(nhanes, gender_factor, age_cat_factor, ed_lvl_factor)

educ_age_gender_nomiss <- na.omit(educ_age_gender, cols = "ed_lvl_factor")

```


```{r vertical bar chart for education level}

ggplot(educ_age_gender_nomiss, 
       aes(x = ed_lvl_factor) ) +
        geom_bar(aes(y = (..count..)/sum(..count..) ), width = .5) +
        labs(x = "highest education level", y = "proportion",
             title = "NHANES 2015-2016 adults age 20+") +
        theme(plot.title = element_text(hjust = 0.5))

```


```{r side by side bar charts with proportions by gender, version 1}

educ_age_gender_nomiss %>% 
        count(gender_factor, ed_lvl_factor) %>%
        group_by(gender_factor) %>%
        mutate(prop = n/sum(n)) %>%
        # the pipe operator (%>%) automatically inserts the dataframe created above
        # as the first argument of the ggplot function below
        ggplot(aes(x = ed_lvl_factor, y = prop, width = .5, fill = gender_factor)) +
        geom_bar(stat = "identity", position = "dodge") +
        labs(y = "proportion", x = "", title = "highest education level NHANES 2015-2016") +
        theme(plot.title = element_text(hjust = 0.5))

```



```{r side by side bar charts with proportions by gender, version 2 horizontal}

educ_age_gender_nomiss %>% 
        count(gender_factor, ed_lvl_factor) %>%
        group_by(gender_factor) %>%
        mutate(prop = n/sum(n))  %>%
        # notice that ggplot uses + to "add" elements to the graph
        # versus using the pipe operator
        ggplot(aes(x = gender_factor, y = prop, width = .5, fill = ed_lvl_factor)) +
        geom_bar(stat = "identity", position = "dodge") + coord_flip() +
        labs(y = "proportion", x = "", title = "highest education level NHANES 2015-2016") +
        theme(plot.title = element_text(hjust = 0.5))

```

```{r stacked bars}

#statisticians find these much less useful!

educ_age_gender_nomiss %>% 
        count(gender_factor, ed_lvl_factor) %>%
        group_by(gender_factor) %>%
        mutate(prop = n/sum(n))  %>%
        ggplot(aes(x = gender_factor, y = prop, width = .5, fill = ed_lvl_factor)) +
        geom_bar(stat = "identity") +
        labs(y = "proportion", x = "", title = "highest education level NHANES 2015-2016") +
        theme(plot.title = element_text(hjust = 0.5))

```

